### 2020-11-30 latex/acrobot.tex
Partial edit of rotations, working on the new proof.

Next, recall that \(R(\pi,\mu_0,I)\) is smooth in all its parameters as well as
being \(O(I^2)\).
Therefore, it is bounded below on the compact set \(\mu_0 \in [\epsilon,\pi]\)
by some value \(R(\pi,\underbar{\mu},I)\). 
Theorem \ref{thm:khalil-perturbation} asserts that there exists some
\(I_3 > 0\) and \(r > 0\) so that, for all \(I \in [-I_3, I_3]\),
\[
    R(\pi,\mu_0,I) \geq R(\pi, \underbar{\mu}, I) > -I^2 r
    .
\]
Furthermore, \(Q(\mu_0)\) is a strictly increasing function, so
for\(\mu_0 \in [\epsilon,\pi[\) we have
\[
    P_\mathcal{O}(\mu_0) > \mu_0 + I K Q(\epsilon) - I^2 r
    .
\]
Choosing \(I^\star \in\, ]0,I_3]\) so that
\(I^\star K Q(\epsilon) - (I^\star)^2 r > 0\)
means that
\( P_\mathcal{O}(\mu_0) > \mu_0\)
for all \(\mu_0 \in [\epsilon,\pi[\).

@@

\subsection{Perturbation Analysis for Rotations}
The rotation region for the nominal pendulum is the set
\[
    \mathcal{R} := \left\{ (q_u,p_u) \in \Gamma \mid E(q_u,p_u) > E(\pi,0)\right\}
    .
\]
Figure \ref{fig:pendulum-level-sets} shows that orbits of the 
pendulum on \(\mathcal{R}\) look like open curves.
Performing a similar process to what we did for oscillations,
we wish to find a new set of coordinates \((\beta,\rho)\)
where \(\rho\) remains constant on level sets of \(E\)
and \(\beta \in \Sone\) is always increasing on \(\mathcal{R}\).

@@

\begin{figure}
    \centering
    %TODO: Add a figure with one rotation mark a distance rho away from the
    %origin along the p-axis. Draw the lower boundary of R and shade the
    %interior, and draw a line to the rotation with angle beta.
    %\includestandalone[]{images/rho_intersection}
    \caption{The domain \(\mathcal{R}\) where a pendulum rotates. The
        pseudo-radius \(\rho\) corresponds to the intersection of an orbit of
        rotation with the \(p_u\)-axis. The pseudo-angle \(\beta\) is taken to
        be clockwise positive, because oscillations of a pendulum move clockwise
        on \(\mathcal{R}\).}
    \label{fig:rho-intersection}
\end{figure}

Note that the boundary of \(\mathcal{R}\) intersects the \(p_u\) axis when 
\(p_u^2 = 60m^2 g l^3\). 
Hence, we must have \(\rho > \sqrt{60 m^2 g l^3}\) if the
orbit has momentum \(p_u > 0\) 
and \(\rho < -\sqrt{60 m^2 g l^3}\) if the orbit has momentum \(p_u < 0\).
The energy level set associated with \(\rho\) is 
\[
    \left\{(q_u,p_u) \in \Gamma \mid E(q_u,p_u) = \frac{\rho^2}{10ml^2}\right\}
    ,
\]
which gives the relationship
\begin{equation}\label{eqn:rotation-pu2}
    \frac{p_u^2}{10m l^2} + 30mgl(1 - c_u) = \frac{\rho^2}{10 ml^2}
    .
\end{equation}
On this level set, \(q_u\) takes all values on \(\Sone\), so our angle of
rotation is uniquely parameterized by \(\beta = q_u\).
Since \(\rho\) does not change sign along the rotation, we have the smooth
relationship
\begin{align}\label{eqn:rotation-Tinv}
    q_u &= \beta
    , \\
    p_u &= \sign{\rho}\sqrt{\rho^2 - 30 m^2 g l^3\left(1 - c_\beta \right)}
    .
\end{align}

@@

Computing the acrobot's constrained dynamics in \((\beta,\rho)\)-coordinates and
setting \(I = 0\) yields the dynamics of the nominal pendulum.
MATLAB evaluates those dynamics as
\begin{align}\label{eqn:acrobot-rot-mu-dot-nom}
    \dot{\rho} &= 0
    , \\
    \label{eqn:acrobot-rot-alpha-dot-nom}
    \dot{\beta} &=  \sign{\rho} 
    \frac{\sqrt{\rho^2 - 30m^2gl^3(1 - c_\beta)}}{5ml^2}
    .
\end{align}
As expected, \(\dot{\beta}\) does not change sign on \(\mathcal{R}\) because the
orbits always flow clockwise.
If \(\rho > 0\), the rotation curve goes from \(\beta = -\pi\) to 
\(\beta = \pi\); 
if \(\rho < 0\), it goes from \(\beta = \pi\) to \(\beta = -\pi\).

We can use \(\beta\) as a new ``time" variable on \(\mathcal{R}\) 
by performing a time transformation \(t = \tau(\beta)\).
This produces the time-scaled system \(\hat{\rho}(\beta) = \rho(\tau(\beta))\)
with dynamics
\[
    \begin{cases}
        \dot{\hat{\rho}} = \frac{\dot{\rho}}{\dot{\beta}}
        , \\
        \hat{\rho}(0) = \rho_0
        .
    \end{cases}
\]
Using the perturbation method, we expand the solution of these dynamics into
\begin{equation}\label{eqn:acrobot-rhohat-approx}
    \hat{\rho}(\beta,\rho_0,I) = \hat{\rho}_0(\beta,\rho_0) +
    I\hat{\rho}_1(\beta,\rho_0) + R(\beta,\rho_0,I)
    ,
\end{equation}
where \(R(\beta,\rho_0,I)\) is smooth and \(O(I^2)\).

From ~\eqref{eqn:khalil-perturbation-nominal} we know
\(\hat{\rho}_0\) is the solution to the nominal system
\[
    \begin{cases} 
        \dot{\hat{\rho}}_0 = 0
        , \\
        \hat{\mu}_0(0) = \rho_0
        ,
    \end{cases}
\]
which means \(\hat{\mu}_0(\beta,\rho_0) \equiv \rho_0\) for all \(\beta\).
Likewise, ~\eqref{eqn:khalil-perturbation-firstorder} 
gives us the following dynamics for \(\hat{\rho}_1\):
\begin{equation}\label{eqn:acrobot-rho1-dot}
  \begin{cases}
    \dot{\hat{\rho}}_1 =
    \frac{5m^2 g l^3 \bar{q}_a \left(
        m^2gl^3\left(18s_\beta^2 + 30c_\beta(1 - c_\beta)\right)
        - c_\beta\rho_0^2
    \right)}{
    |\rho_0|\sqrt{\rho_0^2 - 30m^2gl^3(1 - c_\beta)}
    }
    =: b(\beta,\rho_0)
     , \\
     \hat{\rho}_1(0) = 0
     .
 \end{cases}
\end{equation}
The solution to ~\eqref{eqn:acrobot-rho1-dot} is given by
\[
    \hat{\rho}_1(\beta,\rho_0) = \int \limits_0^\beta b(\sigma,\rho_0)d\sigma
    .
\]

Suppose we initialize the acrobot at \((\beta,\rho) = (0,\rho_0)\).
One full rotation amounts to \(\beta\) traversing \(2\pi\)rad in a clockwise
direction (\ie \(\beta\) going from \(0\) to \(\sign{\rho_0}2\pi\)).
We define the Poincar\'{e} section
\(P_\mathcal{R}(\rho_0) := \hat{\rho}(\sign{\rho_0}2\pi,\rho_0,I)\),
which is the intersection point on the \(p_u\)-axis after a single rotation.
This expands into
\[
    P_\mathcal{R}(\rho_0) = \rho_0 + I \hat{\rho}_1(\sign{\rho_0}2\pi, \rho_0)
      + R(2\pi,\rho_0,I)
    .
\]

The function \(b(\beta,\rho_0)\) is even and is 
\(2\pi\)-periodic in \(\beta\), which implies that
\begin{align*}
    \hat{\rho}_1(\sign{\rho_0}2\pi, \rho_0) &=
    \int \limits_0^{\sign{\rho_0}2\pi} b(\sigma,\rho_0)d\sigma
    , \\
     &=\sign{\rho_0}\int\limits_0^{2\pi}b(\sigma,\rho_0)d\sigma
     , \\
     &= \sign{\rho_0} \hat{\rho}_1(2\pi,\rho_0)
     .
\end{align*}
Furthermore, \(b(\beta,\rho_0)\) is also even in \(\rho_0\), 
so \(\hat{\rho}_1(2\pi,\rho_0) = \hat{\rho}_1(2\pi,\abs{\rho_0})\).
This implies that
\begin{align*}
    \abs{P_\mathcal{R}(\rho_0)} &= \left|\sign{\rho_0}\left(
        \abs{\rho_0} + I \hat{\rho}_1(2\pi,\abs{\rho_0})
        + \sign{\rho_0}R(2\pi,\rho_0,I)\right)\right|
    \\
    &= \left|\abs{\rho_0} + I \hat{\rho}_1(2\pi,\abs{\rho_0})
        + \sign{\rho_0}R(2\pi,\rho_0,I)\right|
    .
\end{align*}

Let now \(\bar{\rho} > \sqrt{60m^2gl^3}\) and take the compact set
\[
    \bar{\mathcal{R}} := \left\{(q_u,p_u) \in \Gamma 
    \mid E(\pi,0) \leq E(q_u,p_u) \leq E(0,\bar{\rho}) \right\}
    .
\]
The (modified) remainder term \(\sign{\rho_0}{R}(\beta,\rho_0,I)\) 
is smooth and \(O(I^2)\), which means
it attains its minimum \(I^2\hat{r}\) somewhere \(\bar{\mathcal{R}}\).
Likewise, \(\hat{\rho}_1(2\pi,\abs{\rho_0})\) also attains its minimum, which we
denote by \(\underbar{\rho}_1\).
We cannot adimensionalize \(b(\beta,\rho_0)\), so refer to one of the 
assumptions of Theorem \ref{thm:acrobot-energy-stabilization}
to suppose  that \(\hat{\rho}_1(2\pi,\rho_0) > 0 \) for all 
\(\rho_0 \in ]\sqrt{60m^2gl^3},\bar{\rho}]\).
In particular, this means \(\underbar{\rho}_1 > 0\).

>>>>

Next, recall that \(R(\pi,\mu_0,I)\) is smooth in all its parameters as well as
being \(O(I^2)\).
Therefore, it is bounded below on the compact set \(\mu_0 \in [\epsilon,\pi]\)
by some value \(R(\pi,\underbar{\mu},I)\). 
Theorem \ref{thm:khalil-perturbation} asserts that there exists some
\(I_3 > 0\) and \(r > 0\) so that, for all \(I \in [-I_3, I_3]\),
\[
    R(\pi,\mu_0,I) \geq R(\pi, \underbar{\mu}, I) > -I^2 r
    .
\]
Furthermore, \(Q(\mu_0)\) is a strictly increasing function, so
for\(\mu_0 \in [\epsilon,\pi[\) we have
\[
    P_\mathcal{O}(\mu_0) > \mu_0 + I K Q(\epsilon) - I^2 r
    .
\]
Choosing \(I^\star \in\, ]0,\min\left\{I_2,I_3\right\}]\) so that
\(I^\star K Q(\epsilon) - (I^\star)^2 r > 0\)
means that
\( P_\mathcal{O}(\mu_0) > \mu_0\)
for all \(\mu_0 \in [\epsilon,\pi[\).

@@

\subsection{Perturbation Analysis for Rotations}
Figure \ref{fig:pendulum-level-sets} reminds us that rotations of the 
nominal pendulum obtained by setting \(I = 0\) look like open curves on some
rotation region \(\mathcal{R} \subset \Gamma\).
One can compute this rotation domain to be the set
\[
    \mathcal{R} := \left\{ (q_u,p_u) \in \Gamma \mid E(q_u,p_u) > E(\pi,0)\right\}
    .
\]
Performing a similar process to what we did for oscillations,
we wish to find a new set of coordinates \((\beta,\rho)\)
where the pseudo-radius \(\rho \in \mathbb{R}\) remains constant on level sets
of \(E\) and the pseudo-angle \(\beta \in \Sone\) is always increasing on
\(\mathcal{R}\).

@@

\begin{figure}
    \centering
    \includestandalone[]{images/rho_intersection}
    \caption{The domain \(\mathcal{R}\) (blue) where a pendulum rotates. The
        pseudo-radius \(\rho\) corresponds to the intersection of an orbit of
        rotation with the \(p_u\)-axis. The pseudo-angle \(\beta\) selects a
        point on the rotation.}
    \label{fig:rho-intersection}
\end{figure}

Note that the boundary of \(\mathcal{R}\) intersects the \(p_u\) axis when 
\(p_u^2 = 60m^2 g l^3\). 
The level set of this boundary is \(E(\pi,0)\), the homoclinic orbit for the
upright equilibrium of the pendulum.
Hence, we must have \(\rho > \sqrt{60 m^2 g l^3}\) if a
rotation has momentum \(p_u > 0\) 
and \(\rho < -\sqrt{60 m^2 g l^3}\) if it has momentum \(p_u < 0\).
The energy level set associated with \(\rho\) is 
\[
    \left\{(q_u,p_u) \in \Gamma \mid E(q_u,p_u) = \frac{\rho^2}{10ml^2}\right\}
    ,
\]
which gives the relationship
\begin{equation}\label{eqn:rotation-pu2}
    \frac{p_u^2}{10m l^2} + 30mgl(1 - c_u) = \frac{\rho^2}{10 ml^2}
    .
\end{equation}
On this level set, \(q_u\) takes all values on \(\Sone\), so our angle of
rotation is uniquely parameterized by \(\beta = q_u\).
Since \(\rho\) does not change sign along the rotation, we have the smooth
relationship
\begin{align}\label{eqn:rotation-Tinv}
    q_u &= \beta
    , \\
    p_u &= \sign{\rho}\sqrt{\rho^2 - 30 m^2 g l^3\left(1 - c_\beta \right)}
    .
\end{align}

@@

Computing the acrobot's constrained dynamics in \((\beta,\rho)\)-coordinates and
setting \(I = 0\) yields the dynamics of the nominal pendulum.
MATLAB evaluates those dynamics as
\begin{align}\label{eqn:acrobot-rot-mu-dot-nom}
    \dot{\rho} &= 0
    , \\
    \label{eqn:acrobot-rot-alpha-dot-nom}
    \dot{\beta} &=  \sign{\rho} 
    \frac{\sqrt{\rho^2 - 30m^2gl^3(1 - c_\beta)}}{5ml^2}
    .
\end{align}
As expected, \(\dot{\beta}\) does not change sign on \(\mathcal{R}\) because the
orbits always flow clockwise.
If \(\rho > 0\), the rotation curve goes from \(\beta = -\pi\) to 
\(\beta = \pi\); 
if \(\rho < 0\), it goes from \(\beta = \pi\) to \(\beta = -\pi\).
By continuity of ~\eqref{eqn:rotation-T}, there exists \(I_1 > 0\) small enough
that \(\dot{\beta}\) is non-zero and does not change sign on \(\mathcal{R}\) for
all \(I \in [-I_1,I_1]\).

Using \(\beta\) as our new time variable produces the time-scaled system
\(\hat{\rho}(\beta) = \rho(\tau(\beta))\) with dynamics
\[
    \begin{cases}
        \dot{\hat{\rho}} = \frac{\dot{\rho}}{\dot{\beta}}
        , \\
        \hat{\rho}(0) = \rho_0
        .
    \end{cases}
\]

In the spirit of perturbation analysis, we expand the time-scaled system 
\(\hat{\rho}(\beta,\rho_0,I)\). 
From ~\eqref{eqn:khalil-perturbation-nominal} we know the nominal system at
\(I = 0\) is
\[
    \begin{cases} 
        \dot{\hat{\rho}}_0 = 0
        , \\
        \hat{\mu}_0(0) = \rho_0
        ,
    \end{cases}
\]
with solution \(\hat{\rho}_0(\beta,\rho_0) \equiv \rho_0\) for all \(\beta\).
Likewise, ~\eqref{eqn:khalil-perturbation-firstorder} 
tells us that \(\hat{\rho}_1(\beta,\rho_0)\) is the solution to the linear
time-varying scalar system
\begin{equation}\label{eqn:acrobot-rho1-dot}
  \begin{cases}
    \dot{\hat{\rho}}_1 =
    \frac{5m^2 g l^3 \bar{q}_a \left(
        m^2gl^3\left(18s_\beta^2 + 30c_\beta(1 - c_\beta)\right)
        - c_\beta\rho_0^2
    \right)}{
    |\rho_0|\sqrt{\rho_0^2 - 30m^2gl^3(1 - c_\beta)}
    }
    =: b(\beta,\rho_0)
     , \\
     \hat{\rho}_1(0) = 0
     ,
 \end{cases}
\end{equation}
whose solution is
\[
    \hat{\rho}_1(\beta,\rho_0) = \int \limits_0^\beta b(\sigma,\rho_0)d\sigma
    .
\]
Therefore, the first-order Taylor approximation of
\(\hat{\rho}(\beta,\rho_0,I)\) is
\begin{equation}\label{eqn:acrobot-rhohat-approx}
    \hat{\rho}(\beta,\rho_0,I) = \hat{\rho}_0(\beta,\rho_0) +
    I\hat{\rho}_1(\beta,\rho_0) + R(\beta,\rho_0,I)
    ,
\end{equation}
where \(R(\beta,\rho_0,I)\) is smooth and \(O(I^2)\).

\subsubsection*{Poincar\'{e} Analysis}

Suppose we initialize the acrobot at \((\beta,\rho) = (0,\rho_0)\).
One full rotation amounts to \(\beta\) traversing \(2\pi\)rad in a clockwise
direction (\ie \(\beta\) going from \(0\) to \(\sign{\rho_0}2\pi\)).
Since \(\dot{\beta}\) does not change sign on \(\mathcal{R}\),
there is a well-defined Poincar\'{e} map describing how the pseudo-radius
\(\rho\) changes each time the orbit \((q(t),p(t))\) hits the \(p_u\) axis, \ie,
every time \(\beta\) changes by \(2\pi\).
We define this Poincar\'{e} map as
\(P_\mathcal{R}(\rho_0) := \hat{\rho}(\sign{\rho_0}2\pi,\rho_0,I)\),
which expands into
\[
    P_\mathcal{R}(\rho_0) = \rho_0 + I \hat{\rho}_1(\sign{\rho_0}2\pi, \rho_0)
    + R(\sign{\rho_0}2\pi,\rho_0,I)
    .
\]

The function \(b(\beta,\rho_0)\) is even and
\(2\pi\)-periodic in \(\beta\), which implies that
\begin{align*}
    \hat{\rho}_1(\sign{\rho_0}2\pi, \rho_0) &=
    \int \limits_0^{\sign{\rho_0}2\pi} b(\sigma,\rho_0)d\sigma
    , \\
     &=\sign{\rho_0}\int\limits_0^{2\pi}b(\sigma,\rho_0)d\sigma
     , \\
     &= \sign{\rho_0} \hat{\rho}_1(2\pi,\rho_0)
     .
\end{align*}
Thus, we have that
\[
    P_\mathcal{R}(\rho_0) = \rho_0 + I\sign{\rho_0}\hat{\rho}_1(2\pi,\rho_0) 
    + R(\sign{\rho_0}2\pi,\rho_0,I)
    .
\]

Note that \(b(\beta,\rho_0)\) is even in \(\rho_0\), which means
\(\hat{\rho}_1(2\pi,\rho_0) = \hat{\rho}_0(2\pi,\abs{\rho_0})\).
Recalling the assumption that there exists \(\bar{\rho} > 0\) so that
\(\int \limits_{0}^{2\pi} b(\sigma,\rho_0)d\sigma > 0\) for all 
\(\rho_0 \in \,]\sqrt{60m^2gl^3},\bar{\rho}]\),
we have that \(\hat{\rho}_1(2\pi,\rho_0) > 0\)
for every \(\rho_0\) with \(\sqrt{60m^2gl^3} < \abs{\rho} \leq \bar{rho}\).
Finally, \(R(\sign{\rho_0}2\pi,\rho_0,I)\) is \(O(I^2)\). 
Defining 
\[
    D := [-\bar{\rho},\sqrt{60m^2gl^3}[ \, \cup \, ]\sqrt{60m^2gl^3},\bar{\rho}]
    ,
\]
these facts all together imply that, for
each \(\rho_0 \in D\),
there exists \(I > 0\) small enough that 
\[
    \begin{cases}
        P_\mathcal{R}(\rho_0) > \rho_0 & \rho_0 > 0
        ,\\
        P_\mathcal{R}(\rho_0) < \rho_0 & \rho_0 < 0
        .\\
    \end{cases}
\]

\subsubsection*{Energy Gain on \(\mathcal{R}\)}

The Poincar\'{e} section \(P_\mathcal{R}\) allows us understand the evolution of
\(\hat{\rho}(\beta,\rho_0,I)\) by studying the evolution of the discrete time
system
\begin{equation}\label{eqn:rhohat-discrete}
    \begin{cases}
        \rho(n+1) := P_\mathcal{R}(\rho(n)) 
        = \rho(n) + I\sign{\rho_0}\hat{\rho}_1(2\pi,\rho(n)) + 
            R(\sign{\rho_0}2\pi,\rho(n),I)
        , \\
        \rho(0) = \rho_0
        .
    \end{cases}
\end{equation}
Tolerating the abuse of notation, \(\rho(n)\) represents the distance along the
\(p_u\)-axis when an orbit of the constrained dynamics intersects the
\(p_u\)-axis for the \(n\)th time, assuming the orbit was initialized at
\((q_u,p_u) = (0,\rho_0)\).

We wish to show \(\abs{\rho(n)}\) eventually reaches \(\bar{\rho}\).
That is, we wish to find some \(I^\star \in \,]0,I_1]\) making
\(\abs{P_\mathcal{R}(\rho_0)} > \abs{\rho_0}\) for all 
\(\rho_0 \in D\)
We will restrict ourselves, without loss of generality, to the case where
\(\rho_0 > 0\).

Recall that \(R(\beta,\rho_0,I)\) is \(O(I^2)\) and smooth in all
its parameters.
Therefore, it is bounded below on the compact set
\(\rho_0 \in [\sqrt{60m^2gl^3},\bar{\rho}]\) by some value
\(R(2\pi,\underbar{\rho},I)\).
Theorem \ref{thm:khalil-perturbation} asserts that there exists some 
\(I_2 > 0\) and \(r > 0\) so that, for all \(I \in [-I_2,I_2]\),
\[
    R(2\pi,\rho_0,I) \geq R(2\pi,\underbar{\rho},I) > -I^2r
    .
\]

%TODO: replace this with an analysis of poincare section
Let now \(\bar{\rho} > \sqrt{60m^2gl^3}\) and take the compact set
\[
    \bar{\mathcal{R}} := \left\{(q_u,p_u) \in \Gamma 
    \mid E(\pi,0) \leq E(q_u,p_u) \leq E(0,\bar{\rho}) \right\}
    .
\]
The (modified) remainder term \(\sign{\rho_0}{R}(\beta,\rho_0,I)\) 
is smooth and \(O(I^2)\), which means
it attains its minimum \(I^2\hat{r}\) somewhere \(\bar{\mathcal{R}}\).
Likewise, \(\hat{\rho}_1(2\pi,\abs{\rho_0})\) also attains its minimum, which we
denote by \(\underbar{\rho}_1\).
We cannot adimensionalize \(b(\beta,\rho_0)\), so refer to one of the 
assumptions of Theorem \ref{thm:acrobot-energy-stabilization}
to suppose  that \(\hat{\rho}_1(2\pi,\rho_0) > 0 \) for all 
\(\rho_0 \in ]\sqrt{60m^2gl^3},\bar{\rho}]\).
In particular, this means \(\underbar{\rho}_1 > 0\).
